Blog post: Different map projections

### 2020-03-17 text/20/map-projections.Rmd
Added mercator projection to map-projections

```{r us-states}
ne_states(country="United States of America", returnclass = "sf") %>%
  filter(postal != "HI" & postal != "AK") -> states
str(states)
```

@@

```{r, echo=FALSE}
projection_string <- "+proj=ortho +lon_0=-80 +lat_0=23.5 +ellps=WGS84 +no_defs"
states %>%
  filter(postal=="CO") -> co
countries %>%
  ggplot() + 
  geom_sf(fill="white") + 
  geom_sf(data=co, color="red", fill="white") +
  coord_sf(crs=projection_string)
```

@@

You can calculate the length of each of these boundaries with a bit of simple math. The radius of the earth is

$r=6,378$ meters

The circumference is

$2\pi r=40,074$ meters

Each degree of latitude would be

$40,074/360=111.3$ meters.

@@

$4*111.3=445.2$ meters

The difference between the northern and southern borders is 4 degrees, so the length of either the eastern or the western borders is

$(4/360)*40,074=445$ meters

The distance from the east to the west border is tricky. At the equator a degree of distance is 111.3 meters, but as you move away from the equator, the lines of latitude converge so a degree of longitude shrinks.

>>>>

```{r us-states}
ne_states(country="United States of America", returnclass = "sf") %>%
  filter(postal == "CO") -> co
str(co)
names(co)
```

@@

```{r, echo=FALSE}
projection_string <- "+proj=ortho +lon_0=-80 +lat_0=23.5 +ellps=WGS84 +no_defs"
countries %>%
  ggplot() + 
  geom_sf(fill="white") + 
  geom_sf(data=co, color="red", fill="white") +
  coord_sf(crs=projection_string)
```

@@

You can calculate the length of each of these boundaries with a bit of simple math. The radius of the earth is 6,378 kilometers. So the circumference is

```{r}
r <- 6378
2*pi*r
```

Each degree of latitude is

```{r}
2*pi*r/360
```

@@

```{r}
2*pi*r*4/360
```

The distance from the east to the west border is tricky. At the equator a degree of distance is 111.3 meters, but as you move away from the equator, the lines of latitude converge so a degree of longitude shrinks. A simple bit of geometry can show you that a single degree of longitude at 37 degrees is

```{r}
2*pi*r*cos(37*pi/180)/360
```

but at 41 degrees, it is

```{r}
2*pi*r*cos(41*pi/180)/360
```

That means that the 9 degrees that you travel from the eastern border to the western border is

```{r}
2*pi*r*cos(37*pi/180)*9/360
```

at 37 degrees, but only

```{r}
2*pi*r*cos(41*pi/180)*9/360
```

So the northern border of Colorado is shorter than the southern border.

And all this time you thought Colorado was a rectangle! It's not. It is not a trapezoid either. It exists on a three dimensional globe and is more complex than any two dimensional geometric object can represent.

You need to make a projection any time you try to represent a three dimensional object like the state of Colorado on a two dimensional medium like a computer screen or a sheet of paper. The most common projection is the Mercator projection. Here is the Mercator projection for the state of Colorado.

```{r mercator}
projection_string <- "+proj=tmrec +lon_0=-80 +lat_0=0 +ellps=WGS84 +no_defs"
projection_string <- "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
countries %>%
  ggplot() + 
  geom_sf(fill="white") + 
  geom_sf(data=co, color="red", fill="white") +
  coord_sf(crs=projection_string, ylim=c(-80,80))
```
